---
output:
  pdf_document: default
  html_document: default
---
<!-- One needs the code below in order to being able of integrating Latex with R -->

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
tex2jax: {
inlineMath: [['$','$'], ['\\(','\\)']],
processEscapes: true},
jax: ["input/TeX","input/MathML","input/AsciiMath","output/CommonHTML"],
extensions: ["tex2jax.js","mml2jax.js","asciimath2jax.js","MathMenu.js","MathZoom.js","AssistiveMML.js", "[Contrib]/a11y/accessibility-menu.js"],
TeX: {
extensions: ["AMSmath.js","AMSsymbols.js","noErrors.js","noUndefined.js"],
equationNumbers: {
autoNumber: "AMS"
}
}
});
</script>

<!-- markdown report begins -->

---
title: "Computational Numerical Statistics  \n Assignment 4"
author: "Migla Miskinyte Reis"
date: "29.10.20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- the function below does not appear in the text as is here to allow us to use some colored text -->

```{r echo=FALSE}
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}
```

<!-- instruction below allows for some extra vertical space-->
&nbsp;

<!-- the font instruction below allows to increase fontsize and the ** gives us boldface -->

<font size="4">

**Exercise 4.** `r colorize("(0.1 points)", "orange")` [**Points: ____**]{style="float:right"}
</font>

Consider \[\textit{I}= \int_{0}^{1} e^{-x}(1-x)^2 \,dx \]

Use m= 10000 and set.seed(1234) in all experiments. 


**4.1** Use the R function integrate() to compute the value of the integral

```{r}
f<-function(x){exp(-x)*(1-x)^2}

f_int<-integrate(f,lower=0,upper=1);f_int

```

**4.2** Now, we will estimate \emph{I} using the antithetic-variable-based Monte Carlo method and report an estimate of variance \textit{I}$_\textit{ant}$ using this method.

```{r}
set.seed(1234) 
m=10000
g <- function(x){exp(-x)*(1-x)^2} 
x=runif(m/2,0,1)
I.hat1=mean(g(x))
I.hat2=mean(g(1-x))
I.hat.a =(I.hat1+I.hat2)/2; I.hat.a
V.a <- 1/m*(1+cor(g(x),g(1-x)))*var(g(x)); V.a
            
```

**4.3** Next we will use MC method to estimate \emph{I} and report \textit{I}$_\textit{MC}$.

```{r}
set.seed(1234)
m=10000
x=runif(m,0,1)
g <- function(x){exp(-x)*(1-x)^2}
I.hat=(1-0)*mean(g(x));I.hat
V.I.hat = var(g(x))/m*(1-0)^2;V.I.hat

```

**4.4** Next, we will estimate percentage of variance reduction when using \textit{I}$_\textit{ant}$ instead of \textit{I}$_\textit{MC}$ using the equation:

\begin{center}

$\textit{percentage}$=$100(1-\displaystyle\frac{V(I_2)}{V(I_1)})$

\end{center}


```{r}

100*(1-V.a/V.I.hat)


```
**4.5** Next, we will use above described variance reduction methods to estimate \textit{I}$_\textit{ant}$ instead of \textit{I}$_\textit{MC}$ using differing m (100, 500, 1000, 5000, 1000).
First, we create a function for antithetic-variable-based methodand estimate \emph{I} and \textit{I}$_\textit{ant}$ for different m.
```{r}
function_ant<-function(m){
  set.seed(1234) 
  f <- function(x){exp(-x)*(1-x)^2}
  x=runif(m/2,0,1)
  I.hat1=mean(g(x))
  I.hat2=mean(g(1-x))
  I.hat.a =(I.hat1+I.hat2)/2; I.hat.a
  V.a <- 1/m*(1+cor(g(x),g(1-x)))*var(g(x)); V.a 
  return(c(I.hat.a,V.a))
}

m100<-function_ant(100)
m500<-function_ant(500)
m1000<-function_ant(1000)
m5000<-function_ant(5000)
m10000<-function_ant(10000)

```

Next, we create a function for usual Monter Carlo variance reduction method and estimate \emph{I} and \textit{I}$_\textit{MC}$ for different m.

```{r}
function_mc<-function(m){
  set.seed(1234) 
  x=runif(m,0,1)
  g <- function(x){exp(-x)*(1-x)^2}
  I.hat=(1-0)*mean(g(x));I.hat
  V.I.hat = var(g(x))/m*(1-0)^2;V.I.hat
  return(c(I.hat.a,V.I.hat))
}

m100m<-function_mc(100)
m500m<-function_mc(500)
m1000m<-function_mc(1000)
m5000m<-function_mc(5000)
m10000m<-function_mc(10000)


```
We also need to estimate percentage for variance reduction for each m between our methods.
```{r}
perc100<-100*(1-m100[2]/m100m[2])
perc500<-100*(1-m500[2]/m500m[2])
perc1000<-100*(1-m1000[2]/m1000m[2])
perc5000<-100*(1-m5000[2]/m5000m[2])
perc10000<-100*(1-m10000[2]/m10000m[2])
```

Finally, we will make a table with our observations. We can see that variance using \textit{I}$_\textit{MC}$ is lower than \textit{I}$_\textit{ant}$ and that overall variance decreases when we increase m.

```{r}
mm<-cbind(m100m,m500m,m1000m,m5000m,m10000m) #created a matrix of vectors for table
row.names(mm)<-c("I_ant Estimate","I_ant Variance")
ma<-cbind(m100,m500,m1000,m5000,m10000)
row.names(ma)<-c("I_MC Estimate","I_MC Variance")
perc<-matrix(c(perc100,perc500,perc1000,perc5000,perc10000),1,5)
row.names(perc)<-c("% var reduction")
table<-rbind(ma,mm,perc)
colnames(table)<-c("100", "500", "1000", "5000", "10000")
knitr::kable(table)
```

<!-- markdown report ends -->