---
output:
  pdf_document: default
  html_document: default
---
<!-- One needs the code below in order to being able of integrating Latex with R -->

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
tex2jax: {
inlineMath: [['$','$'], ['\\(','\\)']],
processEscapes: true},
jax: ["input/TeX","input/MathML","input/AsciiMath","output/CommonHTML"],
extensions: ["tex2jax.js","mml2jax.js","asciimath2jax.js","MathMenu.js","MathZoom.js","AssistiveMML.js", "[Contrib]/a11y/accessibility-menu.js"],
TeX: {
extensions: ["AMSmath.js","AMSsymbols.js","noErrors.js","noUndefined.js"],
equationNumbers: {
autoNumber: "AMS"
}
}
});
</script>

<!-- markdown report begins -->

---
title: "Computational Numerical Statistics  \n Assignment 1"
author: "Migla Miskinyte Reis"
date: "30.09.20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- the function below does not appear in the text as is here to allow us to use some colored text -->

```{r echo=FALSE}
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}
```

<!-- instruction below allows for some extra vertical space-->
&nbsp;

<!-- the font instruction below allows to increase fontsize and the ** gives us boldface -->

<font size="4">

**Exercise 2.** `r colorize("(0.1 points)", "red")` [**Points: ____**]{style="float:right"}
</font>

Consider the random variable $X\sim tn$.
  
(a) Define in R a function called $\texttt{pdf.t}(x,df)$ that corresponds to the p.d.f. of  X and use it to compute $\texttt{pdf.t}(1,3)$. Compare the result with the one given by the built-in R function $\texttt{d.t}()$.

```{r}
pdf.t<-function(x,n){
  gamma((n+1)/2)/(sqrt(n*pi)*gamma(n/2))*(1 + x^2/n)^-((n+1)/2) #my function
}

pdf.t(1,3)                                                    #my calculations

dt(1,3)                                                       #equals to built-in functions
```


(b) Confirm that $\displaystyle\int_{-\infty}^{+\infty}\texttt{pdf.dt}(x)dx=1$. 

```{r}
pdf.t1 = function(x) {
  return(pdf.t(x, 1)) 
  }                              #I fixed my degrees of freedom to 1

integrate(pdf.t1, lower = -Inf, upper=Inf)    #integrate for x


```



(c) Use R built-in functions to construct the usual $tn$ table with $n$ = 1,...,31.

```{r}

alpha<-c(0.1,0.05,0.025,0.01,0.005, 0.001, 0.0005) # make a vector for t distribution probabilities (alpha)
n<-c(seq(1:31))                     # n is my degree of freedom (df)
tn_table<-sapply((1-alpha),function(x) qt(x, df=n)) #using built-in qt quantile function
colnames(tn_table)<-alpha
print(tn_table)

```

<!-- markdown report ends -->