---
output:
  pdf_document: default
  html_document: default
---
<!-- One needs the code below in order to being able of integrating Latex with R -->

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
tex2jax: {
inlineMath: [['$','$'], ['\\(','\\)']],
processEscapes: true},
jax: ["input/TeX","input/MathML","input/AsciiMath","output/CommonHTML"],
extensions: ["tex2jax.js","mml2jax.js","asciimath2jax.js","MathMenu.js","MathZoom.js","AssistiveMML.js", "[Contrib]/a11y/accessibility-menu.js"],
TeX: {
extensions: ["AMSmath.js","AMSsymbols.js","noErrors.js","noUndefined.js"],
equationNumbers: {
autoNumber: "AMS"
}
}
});
</script>



<!-- markdown report begins -->

---
title: "Computational Numerical Statistics  \n Assignment 10"
author: "Migla Miskinyte Reis"
date: "09.12.20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- the function below does not appear in the text as is here to allow us to use some colored text -->

```{r echo=FALSE}
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}
```

<!-- instruction below allows for some extra vertical space-->
&nbsp;

<!-- the font instruction below allows to increase fontsize and the ** gives us boldface -->

<font size="4">

**Exercise 1.** `r colorize("(0.1 points)", "green")` [**Points: ____**]{style="float:right"}
</font>


\\Let $\textit X$ $\sim$N($\mu$, $\sigma^2$), with $\sigma^2$ that is known and p.d.f

\begin{center}

f(x;$\mu$,$\sigma^2$) = $\dfrac{1}{\sqrt{2 \pi \sigma^2}} \exp \left[-\dfrac{(x-\mu)^2}{2\sigma^2} \right]$, where\quad x $\in$ R\quad and $\mu$ $\in$ R

\end{center}


**1.1** Show that this family of distributions belongs to the exponential family. 
\ This is a Gaussian normal distribution and it is known that it can fit a class of exponential families. 
\ For two paramether exponential family we can write a function as:
\begin{center}

f(x;$\theta$) = $\exp\left[\dfrac{x\theta-b(\theta)}{a(\psi)} +c(x,\psi)   \right]$, where $\psi$ is dispersion parameter and $\theta$ is a location parameter, b is a cumulant function.


\end{center}

\ Thus we can re-write our normal distribution: 

\begin{center}

f(x;$\mu$,$\sigma^2$) = $\dfrac{1}{\sqrt{2 \pi \sigma^2}} \exp \left[-\dfrac{(x-\mu)^2}{2\sigma^2} \right]$ = 

\end{center}

by taking a log: 

\begin{center}

$\log f(x;\mu, \sigma^2) = -\frac{1}{2}\log(2\pi\sigma^2) - \frac{(x-\mu)^2}{2 \sigma^2}$

\end{center}

and the exponential:

\begin{center}

 $=$ $\exp\left[-\dfrac{1}{2}\log(2\pi\sigma^2)-\dfrac{(x-\mu)^2}{2\sigma^2}\right]$

\end{center}

\begin{center}

 $=$ $\exp\left[-\dfrac{1}{2}\log(2\pi\sigma^2)-\dfrac{(x^2 -2\mu + \mu^2)}{2\sigma^2}\right]$

\end{center}


\begin{center}

 $=$ $\exp\left[(x\mu - \dfrac{\mu^2}{2})/\sigma^2 +\dfrac{-1}{2}(\dfrac{x^2}{\sigma^2}+log(2\pi\sigma^2))\right]$

\end{center}

Here, we can clearly see that in our equation $x\theta$ is $x\mu$, $b(\theta)$ is $\dfrac{\mu^2}{2}$, $a(\psi)$ is $\sigma^2$ and the rest of the equation is $c(x,\psi)$


**1.2** Is your formulation in 1.1 the canonical form? If not, write it down and identify the canonical link.

In our case we can already see our formulation is the canonical form and that $\mu$ parameter is $\theta$, because $x\theta$ = $x\mu$.  


**1.3** Use the canonical form to show that $\mathbb{E}(X)$ = $\mu$ and V(X) = $\sigma^2$
 
From the canonical function of exponential distribution we have already shown that $x\theta$ = $x\mu$, $b(\theta)$ = $\dfrac{\mu^2}{2}$, $a(\psi)$ = $\sigma^2$. From the properties of exponential family, we know that we can derive moments from canonical function:

\begin{center}

 $E\left[X\right]=b'(\theta)$
 =$ (\dfrac{\theta^2}{2})'=(\dfrac{\mu^2}{2})'=\mu$

\end{center}

And: 


\begin{center}

 $Var\left[X\right]=\psi b''(\theta)$
 =$ \sigma^2(b)''=\sigma^21=\sigma^2$

\end{center}


<!-- markdown report ends -->